### 2018-03-14 submission/manuscript.Rmd
update RF section --ok from josh on understandability

***Random Forest classification analysis.*** 
To classify individuals as having normal colons or tumors, we built Random Forest classification models for each dataset and comparison using taxa with significant ORs (after multiple comparison correction), all taxa, or OTUs. Because no taxa were identified as having a significant OR associated with adenomas using stool samples or tissue samples, classification models based on OR data were not constructed to classify individuals as having normal colons or adenomas. Within the training dataset, 10-fold cross validation (5-fold cross validation for small datasets) was used to build a model that was then evaluated on the testing set. For the models constructed using the taxa with significant ORs, the default mtry setting was used to train the model and this model was tested on the other datasets in the meta-analysis. The reported AUC values are the AUCs for the application of the model on the test sets. For the OTU-based models, the dataset was split into training (80% of samples) and testing (20%) sets and 10-fold cross validation (5-fold cross validation for small datasets) on the training set was used to generate the model for the testing set. The original 80/20 split and fitting was repeated 100 times and the average AUC from these 100 repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of the importance of each taxon to the overall model was used to rank the taxa used in each model. For all models, the default setting used was $\sqrt{p}$, where p is all the variables used in the respective model. Normally, $\sqrt{p}$ has been found to be what is chosen as the ideal mtry after model tuneing [@Breiman2001].

>>>>

***Random Forest classification analysis.*** 
To classify individuals as having normal colons or tumors, we built Random Forest classification models for each dataset and comparison using taxa with significant ORs (after multiple comparison correction), all taxa, or OTUs.  Because no taxa were identified as having a significant OR associated with adenomas using stool samples or tissue samples, classification models based on OR data were not constructed to classify individuals as having normal colons or adenomas. For all models the hyperparameters for ntree was set to 500 and the mtry set to the sqaure root of the number of taxa or OTUs within the model ($\sqrt{total~variables}$). The $\sqrt{total~variables}$ has been found to approximate the optimum mtry chosen after model tuneing [@Breiman2001]. All fecal models were built using a 10-fold cross validation (CV) while tissue models were built using 5-fold CV due to study sample size. One exception to this rule was the Weir study which was built using a 2-fold CV due to extremely low sample number. For significant OR and all taxa data, models were trained on a single study and then tested on the remaining studies with AUCs recorded during both train and testing phases. For the OTU data, 100 10-fold CVs were run to generate a range of AUCs that could be reasonably expected to occur and the average AUC from these 100 repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of the importance of each taxon to the overall model was used to rank the taxa used in each model.

### 2018-03-14 submission/manuscript.tex
update RF section --ok from josh on understandability

\textbf{\emph{Random Forest classification analysis.}} To classify
individuals as having normal colons or tumors, we built Random Forest
classification models for each dataset and comparison using taxa with
significant ORs (after multiple comparison correction), all taxa, or
OTUs. Because no taxa were identified as having a significant OR
associated with adenomas using stool samples or tissue samples,
classification models based on OR data were not constructed to classify
individuals as having normal colons or adenomas. Within the training
dataset, 10-fold cross validation (5-fold cross validation for small
datasets) was used to build a model that was then evaluated on the
testing set. For the models constructed using the taxa with significant
ORs, the default mtry setting was used to train the model and this model
was tested on the other datasets in the meta-analysis. The reported AUC
values are the AUCs for the application of the model on the test sets.
For the OTU-based models, the dataset was split into training (80\% of
samples) and testing (20\%) sets and 10-fold cross validation (5-fold
cross validation for small datasets) on the training set was used to
generate the model for the testing set. The original 80/20 split and
fitting was repeated 100 times and the average AUC from these 100
repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of
the importance of each taxon to the overall model was used to rank the
taxa used in each model. For all models, the default setting used was
\(\sqrt{p}\), where p is all the variables used in the respective model.
Normally, \(\sqrt{p}\) has been found to be what is chosen as the ideal
mtry after model tuneing (52).

>>>>

\textbf{\emph{Random Forest classification analysis.}} To classify
individuals as having normal colons or tumors, we built Random Forest
classification models for each dataset and comparison using taxa with
significant ORs (after multiple comparison correction), all taxa, or
OTUs. Because no taxa were identified as having a significant OR
associated with adenomas using stool samples or tissue samples,
classification models based on OR data were not constructed to classify
individuals as having normal colons or adenomas. For all models the
hyperparameters for ntree was set to 500 and the mtry set to the sqaure
root of the number of taxa or OTUs within the model
(\(\sqrt{total~variables}\)). The \(\sqrt{total~variables}\) has been
found to approximate the optimum mtry chosen after model tuneing (52).
All fecal models were built using a 10-fold cross validation (CV) while
tissue models were built using 5-fold CV due to study sample size. One
exception to this rule was the Weir study which was built using a 2-fold
CV due to extremely low sample number. For significant OR and all taxa
data, models were trained on a single study and then tested on the
remaining studies with AUCs recorded during both train and testing
phases. For the OTU data, 100 10-fold CVs were run to generate a range
of AUCs that could be reasonably expected to occur and the average AUC
from these 100 repeats was reported. The Mean Decrease in Accuracy
(MDA), a measure of the importance of each taxon to the overall model
was used to rank the taxa used in each model.
